---
title: "MC_CellphoneDB_Plots"
output: html_document
date: "2024-07-23"
editor_options: 
  chunk_output_type: console
---

install ktplots if not present:
```{r}
if (!requireNamespace("devtools", quietly = TRUE)) {
    install.packages("devtools")
}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
devtools::install_github("zktuong/ktplots", dependencies = TRUE)
```

packages:
```{r}
library(ktplots)
library(tidyverse)
library(Seurat)
library(ggtext)
library(gridtext)
library(grid)
library(pheatmap)
library(patchwork)
```


call this function (from misc functions file) to over ride pheatmap plotting calls in order to plot formatted text (italics and superscript)
```{r}
override_pheatmap()
```




Load in the data. Files are referenced from analysis_dir. Change to match your folder structure.
```{r}
analysis_dir <- "/R/Shared/"
setwd(analysis_dir)

#change to wherever the seurat object is. Here it is referencing within the analysis_dir.
load("MC_seurat_pub.rbin", verbose = T)


fig_dir <- paste(analysis_dir, "Cpdb_Fig_MC_Fin", sep="/")
if(!dir.exists(fig_dir)){
  dir.create(fig_dir, recursive = T)
}

#immune cells:
pvals <- read.delim("MC_v8_imm_cpdb_out/statistical_analysis_pvalues.txt", check.names = FALSE)

#imm by cohort
pvals_mc <- read.delim("MC_v8_imm_byCohort/imm_mc_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_mc <- read.delim("MC_v8_imm_byCohort/imm_mc_cpdb/statistical_analysis_means.txt", check.names = FALSE)

pvals_cd <- read.delim("MC_v8_imm_byCohort/imm_cd_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_cd <- read.delim("MC_v8_imm_byCohort/imm_cd_cpdb/statistical_analysis_means.txt", check.names = FALSE)

pvals_un <- read.delim("MC_v8_imm_byCohort/imm_un_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_un <- read.delim("MC_v8_imm_byCohort/imm_un_cpdb/statistical_analysis_means.txt", check.names = FALSE)

#all cell groups, by cohort:
pvals_mc_all <- read.delim("MC_v8_all_byCohort/all_mc_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_mc_all <- read.delim("MC_v8_all_byCohort/all_mc_cpdb/statistical_analysis_means.txt", check.names = FALSE)

pvals_cd_all <- read.delim("MC_v8_all_byCohort/all_cd_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_cd_all <- read.delim("MC_v8_all_byCohort/all_cd_cpdb/statistical_analysis_means.txt", check.names = FALSE)

pvals_un_all <- read.delim("MC_v8_all_byCohort/all_un_cpdb/statistical_analysis_pvalues.txt", check.names = FALSE)
means_un_all <- read.delim("MC_v8_all_byCohort/all_un_cpdb/statistical_analysis_means.txt", check.names = FALSE)


#all cell groups:
pvals_all <- read.delim("MC_v8_all_cpdb_out/statistical_analysis_pvalues.txt", check.names = FALSE)
```




#Fig. S18

```{r}
#compatibility between older versions of seurat and newer ones:
seur_har$imm@images <- list()
seur_har$epi@images <- list()
seur_har$str@images <- list()

#epi and stromal idents were from L1, but imm were from L3. Make them all stored in L3 here:
seur_har$epi@meta.data$ident_L3 <- seur_har$epi@meta.data$ident_L1
seur_har$str@meta.data$ident_L3 <- seur_har$str@meta.data$ident_L1

#group all cell classes together into one seurat object
tmp_seur <- merge(seur_har$epi, seur_har$str)
tmp_seur <- merge(tmp_seur, seur_har$imm)
tmp_seur <- NormalizeData(tmp_seur, normalization.method = "LogNormalize", scale.factor = 10000)

seur_all <- tmp_seur

rm(tmp_seur)
```


adjust for formatting:
```{r}
gene_mod <- c("Gzm", "GZMK", "NCR3", "KLRC1", "NR4A", "CD137", "GNLY", "ANXA1", "IL4I1", "HLA-DR", "TNFR", "CD16", "MMP12", "KLF6", "NCAM1", "LTB", "CSF2", "CCL4", "S100B", "PDGFRA")

pvals_mc_all_format <- format_for_cpdb(pvals_mc_all, gene_mod)
means_mc_all_format <- format_for_cpdb(means_mc_all, gene_mod)

pvals_cd_all_format <- format_for_cpdb(pvals_cd_all, gene_mod)
means_cd_all_format <- format_for_cpdb(means_cd_all, gene_mod)

pvals_un_all_format <- format_for_cpdb(pvals_un_all, gene_mod)
means_un_all_format <- format_for_cpdb(means_un_all, gene_mod)

seur_all <- format_seur_for_cpdb(seur_all, "ident_L3", "ident_L3_format", gene_mod)
```

Shorten names to try and make axis more readable....
```{r}

pvals_mc_all_format <- replace_name_cpdb(pvals_mc_all_format, "Enterocyte", "Ent.")
means_mc_all_format <- replace_name_cpdb(means_mc_all_format, "Enterocyte", "Ent.")
pvals_mc_all_format <- replace_name_cpdb(pvals_mc_all_format, "Cycling", "Cyc.")
means_mc_all_format <- replace_name_cpdb(means_mc_all_format, "Cycling", "Cyc.")

pvals_cd_all_format <- replace_name_cpdb(pvals_cd_all_format, "Enterocyte", "Ent.")
means_cd_all_format <- replace_name_cpdb(means_cd_all_format, "Enterocyte", "Ent.")
pvals_cd_all_format <- replace_name_cpdb(pvals_cd_all_format, "Cycling", "Cyc.")
means_cd_all_format <- replace_name_cpdb(means_cd_all_format, "Cycling", "Cyc.")

pvals_un_all_format <- replace_name_cpdb(pvals_un_all_format, "Enterocyte", "Ent.")
means_un_all_format <- replace_name_cpdb(means_un_all_format, "Enterocyte", "Ent.")
pvals_un_all_format <- replace_name_cpdb(pvals_un_all_format, "Cycling", "Cyc.")
means_un_all_format <- replace_name_cpdb(means_un_all_format, "Cycling", "Cyc.")

seur_all@meta.data$ident_L3_format <- gsub(seur_all@meta.data$ident_L3_format, pattern = "Enterocyte", replacement = "Ent.")
seur_all@meta.data$ident_L3_format <- gsub(seur_all@meta.data$ident_L3_format, pattern = "Cycling", replacement = "Cyc.")




mc <- plot_cpdb(
    scdata=seur_all,
    cell_type1="BEST4|DUOX2",
    cell_type2="CD4|CD8",  # this means all cell-types
    celltype_key="ident_L3_format",
    genes=c("CXCL11|CD226"),
    means=means_mc_all_format,
    pvals=pvals_mc_all_format,
    keep_significant_only = FALSE,
    min_interaction_score=0.25
)

cd <- plot_cpdb(
    scdata=seur_all,
    cell_type1="BEST4|DUOX2",
    cell_type2="CD4|CD8",  # this means all cell-types
    celltype_key="ident_L3_format",
    genes=c("CXCL11|CD226"),
    means=means_cd_all_format,
    pvals=pvals_cd_all_format,
    keep_significant_only = FALSE,
    min_interaction_score=0.25
)

un <- plot_cpdb(
    scdata=seur_all,
    cell_type1="BEST4|DUOX2",
    cell_type2="CD4|CD8",  # this means all cell-types
    celltype_key="ident_L3_format",
    genes=c("CXCL11|CD226"),
    means=means_un_all_format,
    pvals=pvals_un_all_format,
    keep_significant_only = FALSE,
    min_interaction_score=0.25
)



mc_pan <- mc +
  theme(legend.position="none",
        axis.text.x = element_text(angle=90, size=7),
        axis.text.x.top=element_markdown()) +
  ggtitle("Microscopic Colitis")
cd_pan <- cd +
  theme(legend.position="none",
        axis.text.x = element_blank()) +
  ggtitle("Chronic Diarrhea")
un_pan <- un +
  theme(legend.position="bottom",
        axis.text.x = element_blank()) +
  ggtitle("Unaffected")




pdf(paste(fig_dir, "SuppFig18_CXCL11_CD226_ByCohort_NameAbbrev.pdf", sep="/"), width = 14, height = 7)
mc_pan / cd_pan / un_pan + plot_layout(guides="collect")
dev.off()

```



Interaction plot:
```{r}
pvals_all_format <- format_for_cpdb(pvals_all, gene_mod)

tmp_plot <- plot_cpdb_heatmap(pvals = pvals_all_format, cellheight = 10, cellwidth = 10)
#legend is too far over -- bring closer to heatmap:
tmp_plot$gtable$widths[4] <- tmp_plot$gtable$widths[4] - unit(100, "bigpts")

pdf(paste(fig_dir, "SuppFig18_IntPlot_all.pdf", sep="/"), width = 15, height = 18)
print(tmp_plot)
dev.off()
```





#Fig. S19

Interaction plot, imm only:
```{r}
pvals_format <- format_for_cpdb(pvals, gene_mod)

tmp_plot <- plot_cpdb_heatmap(pvals = pvals_format, cellheight = 10, cellwidth = 10)
#legend is too far over -- bring closer to heatmap:
tmp_plot$gtable$widths[4] <- tmp_plot$gtable$widths[4] - unit(100, "bigpts")

pdf(paste(fig_dir, "SuppFig19_IntPlot_imm.pdf", sep="/"), width = 13, height = 13)
print(tmp_plot)
dev.off()
```




#Fig. S20
adjust for formatting:
```{r}
pvals_mc_format <- format_for_cpdb(pvals_mc, gene_mod)
means_mc_format <- format_for_cpdb(means_mc, gene_mod)

pvals_cd_format <- format_for_cpdb(pvals_cd, gene_mod)
means_cd_format <- format_for_cpdb(means_cd, gene_mod)

pvals_un_format <- format_for_cpdb(pvals_un, gene_mod)
means_un_format <- format_for_cpdb(means_un, gene_mod)

seur_har$imm <- format_seur_for_cpdb(seur_har$imm, "ident_L3", "ident_L3_format", gene_mod)
```

```{r}
#chemokines
mc_imm <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    gene_family="chemokine",
    means=means_mc_format,
    pvals=pvals_mc_format
    )

cd_imm <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    gene_family="chemokine",
    means=means_cd_format,
    pvals=pvals_cd_format
    )

un_imm <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    gene_family="chemokine",
    means=means_un_format,
    pvals=pvals_un_format
    )


mc_pan <- mc_imm +
  theme(legend.position="none",
        axis.text.x = element_blank()) +
  ggtitle("Microscopic Colitis Chemokines")
cd_pan <- cd_imm +
  theme(legend.position="none",
        axis.text.x = element_blank()) +
  ggtitle("Chronic Diarrhea Chemokines")
un_pan <- un_imm +
  theme(legend.position="bottom",
        axis.text.x = element_blank()) +
  ggtitle("Unaffected Control Chemokines")





#MC Coinhibitory and costimulatory panel
mc_imm_co <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    gene_family=c("Coinhibitory", "Costimulatory"),
    means=means_mc_format,
    pvals=pvals_mc_format
    )

mc_imm_co_pan <- mc_imm_co +
  theme(legend.position="none",
        axis.text.x = element_blank()) +
  ggtitle("Microscopic Colitis Coinhibitory and Costimulatory Factors")




#IFN interaction plots
mc_imm_ifn <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    genes=c("IFNG"),
    means=means_mc_format,
    pvals=pvals_mc_format
    )

cd_imm_ifn <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    genes=c("IFNG"),
    means=means_cd_format,
    pvals=pvals_cd_format
    )

un_imm_ifn <- plot_cpdb(
    scdata=seur_har$imm,
    cell_type1="T CD8 Trm *GZM*<sup>Hi</sup>",
    cell_type2="Macrophage|Monocyte",
    celltype_key="ident_L3_format",
    genes=c("IFNG"),
    means=means_un_format,
    pvals=pvals_un_format,
    keep_significant_only = FALSE 
    )
#unaffected has no significant int.

mc_ifn_pan <- mc_imm_ifn +
  theme(legend.position="none",
        axis.text.x = element_text(angle=60),
        axis.text.x.top=element_markdown()) +
  ggtitle("Microscopic Colitis IFNG")

cd_ifn_pan <- cd_imm_ifn +
  theme(legend.position="none",
        axis.text.x = element_blank()) +
  ggtitle("Chronic Diarrhea IFNG")




pdf(paste(fig_dir, "SuppFig20_MC.pdf", sep="/"), width = 12, height = 18)
mc_ifn_pan / cd_ifn_pan / mc_pan / cd_pan / un_pan + mc_imm_co_pan + plot_layout(guides="collect", heights=c(0.5, 0.5, 2, 1, 1, 8))
dev.off()
```










